Analysis-ready VCF at Biobank scale using Zarr

### 2024-05-30 paper.tex
Clear up a minor comment from Shadi

\subsection{Subsetting the genotype matrix}
As datasets grow ever larger, the ability to efficiently access subsets 
of the data becomes more and more important. VCF/BCF achieve efficient 
access to the data for genomic ranges 
by compressing blocks of adjacent records using \texttt{bgzip},
and storing secondary indexes alongside the original 
files with a conventional suffix~\citep{li2011tabix}. 
Thus, for a given range query we 
decompress only the necessary blocks and can quickly access
the required records. 
Methods like Savvy and XSI largely follow the same principles [TODO check and 
be more concrete]. The row-wise nature of these methods, however, mean 
that we cannot efficiently subset \emph{by sample}
(e.g., to calculate statistics within a particular cohort). In the extreme
case, if we want to access only the genotypes for a single sample
we must still retrieve and decompress the entire dataset.

\begin{figure}
\includegraphics{figures/subset-matrix-compute}
\caption{Compute performance on subsets of the matrix.
Total CPU time required to run the af-dist calculation for
a subset of 10,000 variants $\times$ 10 samples from the middle of the matrix
for the data in Fig~\ref{fig-data-storage}.
The \texttt{genozip} and \texttt{bcftools} pipelines involve
multiple commands required to correctly calculate the AF INFO field
required by \texttt{bcftools +af-dist}. See the Methods for full details
on the steps performed.
\label{fig-subset-matrix-compute}}
\end{figure}

>>>>

\subsection{Subsetting the genotype matrix}
As datasets grow ever larger, the ability to efficiently access subsets 
of the data becomes increasingly important. VCF/BCF achieve efficient 
access to the data for genomic ranges 
by compressing blocks of adjacent records using \texttt{bgzip},
and storing secondary indexes alongside the original 
files with a conventional suffix~\citep{li2011tabix}. 
Thus, for a given range query we 
decompress only the necessary blocks and can quickly access
the required records. 
Methods like Savvy and XSI largely follow the same principles [TODO check and 
be more concrete]. The row-wise nature of these methods, however, mean 
that we cannot efficiently subset \emph{by sample}
(e.g., to calculate statistics within a particular cohort). In the extreme
case, if we want to access only the genotypes for a single sample
we must still retrieve and decompress the entire dataset.

\begin{figure}
\includegraphics{figures/subset-matrix-compute}
\caption{Compute performance on subsets of the matrix.
Total CPU time required to run the af-dist calculation for
a contiguous subset of 10,000 variants $\times$ 10 samples 
from the middle of the matrix
for the data in Fig~\ref{fig-data-storage}.
The \texttt{genozip} and \texttt{bcftools} pipelines involve
multiple commands required to correctly calculate the AF INFO field
required by \texttt{bcftools +af-dist}. See the Methods for full details
on the steps performed.
\label{fig-subset-matrix-compute}}
\end{figure}
